
---
title: 'IeDEA MR190: DTG rollout <br> 03_CROI2023_Poster_Table1_Fig1-3'
author: "Elizabeth Zaniewski, Veronika Whitesell, Marie Ballif"
date: "`r Sys.Date()`"
output:
  rmdformats::robobook:
    code_folding: show
    self_contained: true
    highlight: pygments
    keep_md: no
    toc_depth: 3
editor_options: 
  chunk_output_type: inline
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "05_html") })
---

```{r r-setup, include = FALSE}
options(scipen = 999)
options(max.print = "75")
set.seed(12345)

library(pacman)
p_load(
  kableExtra, tidyverse, dplyr, Hmisc,
  scales, ggplot2,
  fst, data.table, psych, readxl,
  sjmisc, flextable
)

import::from("psych", "geometric.mean")
import::from("sjmisc", "frq")
import::from("gmodels", "CrossTable")
```

```{r conflicts, include = FALSE}
conflicted::conflict_scout()
```

```{r knit-setup, include = FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

knitr::opts_chunk$set(
  cache = FALSE,
  prompt = FALSE,
  tidy = FALSE,
  comment = NA,
  message = FALSE,
  warning = FALSE,
  echo = TRUE
)

knitr::opts_knit$set(width = 75)
```



---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- ----------------------------------------------------- -->


```{r include=FALSE}
#Clear existing data and graphics
rm(list=ls())
graphics.off()
```

<!-- ----------------------------------------------------- -->


# CROI Poster presentation

https://www.croiconference.org/wp-content/uploads/sites/2/posters/2023/MR190_CROI2023_poster_14Feb2023-133208724161778484.pdf


TRANSITION TO DOLUTEGRAVIR-BASED ART IN LOW- AND MIDDLE-INCOME COUNTRIES IN IEDEA


<!-- ----------------------------------------------------- -->

# Table 1: Type of DTG policy initiative  

Table 1: Type of initiative used to rollout DTG regimens


```{r include=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(sjmisc)
library(gmodels)
library(Hmisc)
library(readxl)
library(writexl)
library(table1)
library(gtsummary)
library(stats)
library(rstatix)

import::from("sjmisc", "frq")
import::from("gmodels", "CrossTable")


import::from("psych", "geometric.mean")
import::from("sjmisc", "frq")
import::from("gmodels", "CrossTable")
```


## Load wrangled data
```{r}
df <- readRDS("03_Output/Data/data_190_variables.Rda") 
```

Number of unique regions, countries and sites
```{r echo=FALSE}
# find unique number of regions and countries
df %>%
  select(region, country, record_id) %>%
  sapply(n_distinct)
```

## Re-wrangle data

HIV prevalence category headings

```{r}
df$Prev_cat.factor <- factor(df$Prev_cat, levels = c( "Low","Medium" ,"High"),
                        labels = c("Low", "Medium", "High" ))

label(df$dtg_rollout_type.factor)="Initiative type"

```


## Create table

Table 1: Type of initiative used to rollout DTG regimens

## Include function used to define table
```{r}


my.render.cat <- function(x) {
  c("", sapply(stats.default(x), function(y) with(y,
                                                  sprintf("%0.0f%% (%d)", PCT, FREQ))))
}

```

## Final Table 1 

(with Total percent and number)
```{r echo=FALSE}

table1 <- table1::table1(~ dtg_rollout_type.factor | Prev_cat.factor, data = df, overall=c(right="Total"), render.categorical = my.render.cat)

table1

table1 <- as.data.frame(table1)


```


Remove "Total" column and export to Excel

```{r echo=FALSE}
table1b <- table1::table1(~ dtg_rollout_type.factor | Prev_cat.factor, data = df, overall=c(right="Total"), render.categorical = my.render.cat)

table1b <- as.data.frame(table1b)

# remove Total column
table1b <- table1b[, 1:4]

Final_table1 <- as.data.frame(table1b)

html_table1 <- qflextable(Final_table1)


#table1b

```


```{r}
html_table1
```


Table 1: Type of initiative used to rollout DTG regimens



## Output Table 1

```{r}
   
    # Print plot to a png file
write_xlsx(Final_table1, "03_Output/CROI_poster/Table_1.xlsx")
    
```

```{r include=FALSE}

rm(df)
rm("table1", "table1b", "html_table1")

```


<!-- ----------------------------------------------------- -->

# Figure 1: DTG rollout or planned rollout   

Figure 1: Transition to DTG regimens by HIV prevalence category and treatment line

For each first-, second-, third-line and across all three lines by HIV prev cat


## Load wrangled data
```{r}
df <- readRDS("03_Output/Data/data_190_variables.Rda") 
```

Number of unique regions, countries and sites
```{r echo=FALSE}
# find unique number of regions and countries
df %>%
  select(region, country, record_id) %>%
  sapply(n_distinct)
```



## Re-wrangle data: overall

Create numerators overall

Rollout or planned rollout
```{r}

# Outcome ab = rolled out

# dtg1 
dtg1_ab_all <- df %>%
   summarise(
    numer = sum(dtg1),
    denom = n()) %>%
  mutate(line = "dtg1") %>%
  mutate(outcome = "ab") %>%
  mutate(per = numer/denom *100) %>%
  mutate(value = round(per, 1)) %>%
  select(-c(per))

# dtg2 by Prev_cat
dtg2_ab_all <- df %>%
  summarise(
    numer = sum(dtg2),
    denom = n()) %>%
  mutate(line = "dtg2") %>%
  mutate(outcome = "ab") %>%
  mutate(per = numer/denom *100) %>%
  mutate(value = round(per, 1)) %>%
  select(-c(per))

# dtg3 by Prev_cat
dtg3_ab_all <- df %>%
   summarise(
    numer = sum(dtg3),
    denom = n()) %>%
  mutate(line = "dtg3") %>%
  mutate(outcome = "ab") %>%
  mutate(per = numer/denom *100) %>%
  mutate(value = round(per, 1)) %>%
  select(-c(per))

```

Create numerators overall

No rollout

```{r}
# Outcome c = no planned rollout or unknown

# dtg1 
dtg1_c_all <- df %>%
  summarise(
    denom = n(),
    numer_ab = sum(dtg1),
    numer = denom - numer_ab) %>%
  mutate(line = "dtg1") %>%
  mutate(outcome = "c") %>%
  mutate(per = numer/denom *100) %>%
  mutate(value = round(per, 1)) %>%
  select(-c(per, numer_ab))

# dtg2 by Prev_cat
dtg2_c_all <- df %>%
  summarise(
    denom = n(),
    numer_ab = sum(dtg2),
    numer = denom - numer_ab) %>%
  mutate(line = "dtg2") %>%
  mutate(outcome = "c") %>%
  mutate(per = numer/denom *100) %>%
  mutate(value = round(per, 1)) %>%
  select(-c(per, numer_ab))

# dtg3 by Prev_cat
dtg3_c_all <- df %>%
  summarise(
    denom = n(),
    numer_ab = sum(dtg3),
    numer = denom - numer_ab) %>%
  mutate(line = "dtg3") %>%
  mutate(outcome = "c") %>%
  mutate(per = numer/denom *100) %>%
  mutate(value = round(per, 1)) %>%
  select(-c(per, numer_ab))
```


Append files
```{r}
dtg_all <- rbind(rbind(dtg1_ab_all, dtg1_c_all, dtg2_ab_all, dtg2_c_all, dtg3_ab_all, dtg3_c_all))
dtg_all <- dtg_all %>%
  mutate(Prev_cat = "All")
```

Rollout or planned rollout all three lines
```{r}
# Roll out of DTG for all three lines
dtg123_ab_all <- df %>%
    summarise(
    denom = n(),
    numer = sum(dtg123)) %>%
  mutate(line = "dtg123") %>%
  mutate(outcome = "ab") %>%
  mutate(per = numer/denom *100) %>%
  mutate(value = round(per, 1)) %>%
  select(-c(per))
```

No rollout all three lines
```{r}

# All for no roll out of DTG for all three  lines
dtg123_c_all <- df %>%
    summarise(
    denom = n(),
    numer_ab = sum(dtg123, na.rm =TRUE),
    numer = denom - numer_ab) %>%
  mutate(line = "dtg123") %>%
  mutate(outcome = "c") %>%
  mutate(per = numer/denom *100) %>%
  mutate(value = round(per, 1)) %>%
  select(-c(per, numer_ab))

```

Append files
```{r}
dtg123_all <- rbind(dtg123_ab_all, dtg123_c_all)
dtg123_all <- dtg123_all %>%
  mutate(Prev_cat = "All")
```



## Re-wrangle data: by HIV prevalence

We linked site data to national HIV prevalence estimates for 2020, obtained from UNAIDS1 and other sources, and categorized as low (<1%), medium (1-4.9%), or high (≥5%) HIV prevalence.

Rollout or planned rollout
```{r}

# Outcome ab = rolled out

# dtg1 by Prev_cat
dtg1_ab <- df %>%
  group_by(Prev_cat) %>% 
  summarise(
    numer = sum(dtg1),
    denom = n()) %>%
  mutate(line = "dtg1") %>%
  mutate(outcome = "ab") %>%
  mutate(per = numer/denom *100) %>%
  mutate(value = round(per, 1)) %>%
  select(-c(per))

# dtg2 by Prev_cat
dtg2_ab <- df %>%
  group_by(Prev_cat) %>% 
  summarise(
    numer = sum(dtg2),
    denom = n()) %>%
  mutate(line = "dtg2") %>%
  mutate(outcome = "ab") %>%
  mutate(per = numer/denom *100) %>%
  mutate(value = round(per, 1)) %>%
  select(-c(per))

# dtg3 by Prev_cat
dtg3_ab <- df %>%
  group_by(Prev_cat) %>% 
  summarise(
    numer = sum(dtg3),
    denom = n()) %>%
  mutate(line = "dtg3") %>%
  mutate(outcome = "ab") %>%
  mutate(per = numer/denom *100) %>%
  mutate(value = round(per, 1)) %>%
  select(-c(per))

```

No rollout
```{r}
# Outcome c = no planned rollout or unknown

# dtg1 by Prev_cat
dtg1_c <- df %>%
  group_by(Prev_cat) %>%
  summarise(
    denom = n(),
    numer_ab = sum(dtg1),
    numer = denom - numer_ab) %>%
  mutate(line = "dtg1") %>%
  mutate(outcome = "c") %>%
  mutate(per = numer/denom *100) %>%
  mutate(value = round(per, 1)) %>%
  select(-c(per, numer_ab))

# dtg2 by Prev_cat
dtg2_c <- df %>%
  group_by(Prev_cat) %>%
  summarise(
    denom = n(),
    numer_ab = sum(dtg2),
    numer = denom - numer_ab) %>%
  mutate(line = "dtg2") %>%
  mutate(outcome = "c") %>%
  mutate(per = numer/denom *100) %>%
  mutate(value = round(per, 1)) %>%
  select(-c(per, numer_ab))

# dtg3 by Prev_cat
dtg3_c <- df %>%
  group_by(Prev_cat) %>%
  summarise(
    denom = n(),
    numer_ab = sum(dtg3),
    numer = denom - numer_ab) %>%
  mutate(line = "dtg3") %>%
  mutate(outcome = "c") %>%
  mutate(per = numer/denom *100) %>%
  mutate(value = round(per, 1)) %>%
  select(-c(per, numer_ab))
```

Append files
```{r}
dtg_HIVprev <- rbind(rbind(dtg1_ab, dtg1_c, dtg2_ab, dtg2_c, dtg3_ab, dtg3_c))

```

Rollout all three lines

```{r}
dtg123_ab <- df %>%
  group_by(Prev_cat) %>%
  summarise(
    denom = n(),
    numer = sum(dtg123)) %>%
  mutate(line = "dtg123") %>%
  mutate(outcome = "ab") %>%
  mutate(per = numer/denom *100) %>%
  mutate(value = round(per, 1)) %>%
  select(-c(per))

```

No rollout all three lines

```{r}
dtg123_c <- df %>%
  group_by(Prev_cat) %>%
  summarise(
    denom = n(),
    numer_ab = sum(dtg123, na.rm =TRUE),
    numer = denom - numer_ab) %>%
  mutate(line = "dtg123") %>%
  mutate(outcome = "c") %>%
  mutate(per = numer/denom *100) %>%
  mutate(value = round(per, 1)) %>%
  select(-c(per, numer_ab))
```

Append files

```{r}
dtg123_HIVprev<- rbind(dtg123_ab, dtg123_c)
#dim(dtg123_HIVprev)  
```



## Append files overall and HIV prevalence
```{r}
Prev_cat_all <- rbind(dtg_all, dtg123_all, dtg_HIVprev, dtg123_HIVprev)
#dim(Prev_cat_all)  
```


## Final Figure 1

Figure 1: Transition to DTG regimens by HIV prevalence category and treatment line

Define plot labels

```{r}
data1 <- Prev_cat_all


data1 <- data1 %>%
  mutate(text = ifelse(data1$outcome =="ab", round(value), "")) %>%
  mutate(text2 = "%")

data1$text3 <- paste(data1$text, data1$text2, sep="")

data1 <- data1 %>%
  mutate(text4 = ifelse(data1$outcome =="ab", text3, "")) 

```

Define plot factors

```{r}
data1$outcome <- factor(data1$outcome, levels = c("ab","c"),
                       labels = c("Rolled out or planned rollout", "No rollout or unknown")) 
  
data1$Prev_cat <- factor(data1$Prev_cat, levels = c( "Low","Medium" ,"High", "All"),
            labels = c("Low HIV prevalence \n (n=35)", "Medium HIV prevalence \n (n=83)", "High HIV prevalence \n (n=57)", "Overall \n (n=175)"))
  
  
data1$line <- factor(data1$line, levels = c("dtg1","dtg2", "dtg3", "dtg123"),
                          labels = c("1st line","2nd line", "3rd line", "All lines"))
  
fill <- c("#3F6D9B", "#a94c4c")
   
```



HIV prev cat only (legend on bottom)

Final Figure 1 in CROI Poster
```{r}
 # stacked bar chart (without overall plot; don't need overall as included in poster text)
    data2 <- data1[-c(1:8), ]

 # legend on bottom
    Final_figure1 <- ggplot(data=data2, aes(x=line, y=value, fill=outcome, label = text4))  + 
      geom_bar(position = position_stack(reverse = TRUE), stat="identity") +
      facet_grid(. ~ Prev_cat) +
      theme(strip.text = element_text(size=11)) +
      theme(legend.position="bottom", 
            legend.box.background = element_rect(color = "black"),
            legend.direction="horizontal", 
            legend.margin=margin(0,0,0,0),
            legend.box.margin = margin(-5, -5, -5, -5), 
            legend.title = element_blank(),
            legend.text = element_text(size=11)) +
      labs(x="", y="Percentage (%)") +
      theme(axis.text.x = element_text(angle =90, hjust =1, vjust =0.5, size =11)) +
      theme(axis.text.y = element_text(size =11)) +
      theme(axis.title.y = element_text(size=11)) +
      scale_y_continuous(n.breaks = 5) +
      scale_fill_manual(labels = ~ stringr::str_wrap(.x, width = 15), values = fill)  +
      scale_color_manual(labels = ~ stringr::str_wrap(.x, width = 15), values = fill) +
      guides(fill = guide_legend(reverse=FALSE, bycol=TRUE)) +
      geom_text(size = 3.5, colour = "white", fontface = "bold", position = position_stack(reverse = TRUE, vjust = 0.8))
    
   
    
```

```{r}
 Final_figure1
```


Figure 1: Transition to DTG regimens by HIV prevalence category and treatment line



## Output Figure 1

```{r}
   
    # Print plot to a png file
    ggsave("03_Output/CROI_poster/Figure1_DTG_rollout_by_HIVprev.png", width = 6, height =3)
    
```



```{r include=FALSE}
rm(df)
rm(list=ls(pattern="dtg"))
rm(list=ls(pattern="data"))
rm(Prev_cat_all)

```


<!-- ----------------------------------------------------- -->

# Figure 2: VL monitoring and DR testing availability   

Figure 2: VL monitoring and DR testing availability and use at switch to DTG regimens by HIV prevalence category


## Load wrangled data
```{r}
df <- readRDS("03_Output/Data/data_190_variables.Rda") 
```

Number of unique regions, countries and sites
```{r echo=FALSE}
# find unique number of regions and countries
df %>%
  select(region, country, record_id) %>%
  sapply(n_distinct)
```


## Re-wrangle data: for availability

```{r}

# create new variables for availability

# VL

df$pcr_vl_yn <- ifelse(df$pcr_vl_na =="1", 0, 1)
CrossTable(df$pcr_vl_yn, df$pcr_vl_na, prop.c = FALSE, prop.t = FALSE, prop.chisq = FALSE)

#frq(df,pcr_vl_clinic)
#frq(df,pcr_vl_samefacility)
#frq(df,pcr_vl_offsite)
CrossTable(df$Prev_cat, df$pcr_vl_yn,
           prop.c = FALSE, prop.t = FALSE, prop.chisq = FALSE)


# DR
df$dr_yn <- ifelse(df$dr_na ==1, 0, 1)
CrossTable(df$dr_yn, df$dr_na, prop.c = FALSE, prop.t = FALSE, prop.chisq = FALSE)

#frq(df,dr_clinic)
#frq(df,dr_samefacility)
#frq(df,dr_offsite)
CrossTable(df$Prev_cat, df$dr_yn,
           prop.c = FALSE, prop.t = FALSE, prop.chisq = FALSE)


```




## Re-wrangle data: for availability and use at switch

Viral load testing

```{r}

# create 5 categories
# 1 Available and used for switch
# 2 Available but not used for switch
# 3 (Available but not known if used for switch --NA)
# 4 Not available
# 5 Not available, but used for switch


# VL
df1_VL <- df %>%
    mutate(testVL_cat1 = ifelse(pcr_vl_yn ==1 & (dtg_trans_vl_yn ==1), 1, 0)) %>%  # VL available and used for switch
    mutate(testVL_cat2 = ifelse(pcr_vl_yn ==1 & (dtg_trans_vl_yn ==0), 1, 0)) %>%   # VL available and not used for switch
    mutate(testVL_cat3 = ifelse(pcr_vl_yn ==1 & is.na(dtg_trans_vl_yn), 1, 0)) %>%   # VL available but switch is NA
    mutate(testVL_cat4 = ifelse(pcr_vl_yn ==0 & (dtg_trans_vl_yn ==0), 1, 0)) %>%    # VL not available and not used for switch
    mutate(testVL_cat5 = ifelse(pcr_vl_yn ==0 & (dtg_trans_vl_yn ==1), 1, 0)) %>%   # VL not available but used for switch 
    mutate(testVL_cat6 = ifelse(pcr_vl_yn ==0 & is.na(dtg_trans_vl_yn), 1, 0)) %>%   # VL not available but switch is NA
    mutate(outcome = ifelse(testVL_cat1 ==1, "1", "0")) %>%
    mutate(outcome = ifelse(testVL_cat2 ==1, "2", outcome)) %>%
    mutate(outcome = ifelse(testVL_cat3 ==1, "3", outcome)) %>%
    mutate(outcome = ifelse(pcr_vl_yn ==0, "4", outcome)) %>%
    mutate(test = "VL") %>%
    mutate(temp1 =1) 

# 1 = available, used at switch
# 2 = available, not used at switch
# 3 = available, unknown if used at switch
# 4 = not available

frq(df1_VL$outcome)

df2_VL <- df1_VL %>%
  group_by(Prev_cat) %>%
  mutate(denom = sum(temp1))

df3_VL <- df2_VL %>%
  group_by(Prev_cat, outcome) %>%
  mutate(numer = sum(temp1)) %>%
  mutate(per = numer/denom *100) %>%
  mutate(value = round(per, 0)) %>%
  select(-c(per))

df4_VL <- df3_VL %>%
  group_by(Prev_cat, test, numer, denom, value, outcome) %>%
  summarise()


# Overall VL

df3_VL_all <- df2_VL %>%
  mutate(denom = sum(temp1)) %>%
  mutate(Prev_cat = "All") 

df3_VL_all <- df3_VL_all %>%
  group_by(outcome) %>%
  mutate(numer = sum(temp1)) %>%
  mutate(per = numer/denom *100) %>%
  mutate(value = round(per, 0)) %>%
  select(-c(per))

df4_VL_all <- df3_VL_all %>%
  group_by(Prev_cat, test, numer, denom, value, outcome) %>%
  summarise()
```


Drug resistance testing


```{r}

# DR 
df1_DR <- df %>%
  mutate(testDR_cat1 = ifelse(dr_na ==0 & dtg_genotyping ==1, 1, 0)) %>%  # DR available and used for switch
  mutate(testDR_cat2 = ifelse(dr_na ==0 & dtg_genotyping ==0, 1, 0)) %>%   # DR available and not used for switch
  mutate(testDR_cat3 = ifelse(dr_na ==0 & is.na(dtg_genotyping), 1, 0)) %>%   # DR available but switch is NA
  mutate(testDR_cat4 = ifelse(dr_na ==1 & dtg_genotyping ==0, 1, 0)) %>%    # DR not available and not used for switch
  mutate(testDR_cat5 = ifelse(dr_na ==1 & dtg_genotyping ==1, 1, 0)) %>%   # DR not available but used for switch 
  mutate(testDR_cat6 = ifelse(dr_na ==1 & is.na(dtg_genotyping), 1, 0)) %>%   # DR not available but switch is NA
  mutate(outcome = ifelse(testDR_cat1 ==1, "1", 0)) %>%
  mutate(outcome = ifelse(testDR_cat2 ==1, "2", outcome)) %>%
  mutate(outcome = ifelse(testDR_cat3 ==1, "3", outcome)) %>%
  mutate(outcome = ifelse(dr_yn ==0, "4", outcome)) %>%
  mutate(test = "DR") %>%
  mutate(temp1 =1) 

# 1 = available, used at switch
# 2 = available, not used at switch
# 3 = available, unknown if used at switch
# 4 = not available

frq(df1_DR$outcome)


df2_DR <- df1_DR %>%
  group_by(Prev_cat) %>%
  mutate(denom = sum(temp1))


df3_DR <- df2_DR %>%
  group_by(Prev_cat, outcome) %>%
  mutate(numer = sum(temp1)) %>%
  mutate(per = numer/denom *100) %>%
  mutate(value = round(per, 0)) %>%
  select(-c(per)) 

  
df4_DR <- df3_DR %>%
  group_by(Prev_cat, test, numer, denom, value, outcome) %>%
  summarise()


# Overall DR

df3_DR_all <- df2_DR %>%
  mutate(denom = sum(temp1)) %>%
  mutate(Prev_cat = "All") 

df3_DR_all <- df3_DR_all %>%
  group_by(outcome) %>%
  mutate(numer = sum(temp1)) %>%
  mutate(per = numer/denom *100) %>%
  mutate(value = round(per, 0)) %>%
  select(-c(per))

df4_DR_all <- df3_DR_all %>%
  group_by(Prev_cat, test, numer, denom, value, outcome) %>%
  summarise()
```

## Append files

```{r}
data3 <- rbind(df4_VL, df4_VL_all, df4_DR, df4_DR_all)
```



## Final Figure 2

Figure 2: VL monitoring and DR testing availability and use at switch to DTG regimens by HIV prevalence category

Define plot labels

```{r}

data3 <- data3 %>%
  mutate(text1 = value) %>%
  mutate(text2 = "%") %>%
  mutate(text3 = paste(text1, text2, sep="")) %>%
  mutate(text4 = ifelse(outcome =="1", text3, ""))
```

Define plot factors

```{r}
data3 <- data3 %>% filter(Prev_cat != "All")

data3$test <- factor(data3$test, levels = c("VL", "DR"),
              labels = c("VL", "DR"))
  
  
data3$outcome <- factor(data3$outcome, levels = c(1, 2, 3, 4),
                       labels = c("Available, used at switch", "Available, not used at switch", "Available, unknown if used at switch", "Not available")) 


data3$Prev_cat <- factor(data3$Prev_cat, levels = c( "Low","Medium", "High"),
                        labels = c("Low \n  HIV prevalence \n (n=35)", "Medium \n HIV prevalence \n (n=83)", "High \n HIV prevalence \n (n=57)"))


fill <- c("royalblue4", "deepskyblue3", "lightskyblue1", "#a94c4c")


   
```



Viral load and drug resistance (legend on bottom)
```{r}

# stacked bar chart with legend on bottom  
Final_figure2 <- ggplot(data=data3, aes(x=test, y=value, fill=outcome, label = text4))  + 
  geom_bar(position = position_stack(reverse = TRUE), stat="identity") +
  facet_grid(. ~ Prev_cat) +
  theme(strip.text = element_text(size=14)) +
  theme(legend.position="bottom", 
        legend.box.background = element_rect(color = "black"),
        legend.direction="horizontal", 
        legend.margin=margin(0,0,0,0),
        legend.box.margin = margin(-5, -5, -5, -5), 
        legend.title = element_blank(),
        legend.key.size = unit(0.7, "cm"), 
        legend.text = element_text(size=14)) +
  labs(x="", y="Percentage (%)") +
  theme(axis.text.x = element_text(angle =90, hjust =1, vjust =0.5, size =14)) +
  theme(axis.text.y = element_text(size =14)) +
  theme(axis.title.y = element_text(size=14)) +
  scale_y_continuous(n.breaks = 5) +
  scale_fill_manual(labels = ~ stringr::str_wrap(.x, width = 12), values = fill)  +
  scale_color_manual(labels = ~ stringr::str_wrap(.x, width = 12), values = fill) +
  guides(fill = guide_legend(reverse=FALSE, byrow=TRUE)) +
  geom_text(size = 5, colour = "white", fontface = "bold", position = position_stack(reverse = TRUE, vjust = 0.8))


```

```{r}
Final_figure2
```


Figure 2: VL monitoring and DR testing availability and use at switch to DTG regimens by HIV prevalence category



## Output Figure 2

```{r}
   
    # Print plot to a png file
ggsave("03_Output/CROI_poster/Figure2_VL_DR_availability.png", width = 7, height =4.5)
    
```


```{r include=FALSE}

rm(list=ls(pattern="df"))
rm(list=ls(pattern="data"))

```



<!-- ----------------------------------------------------- -->

# Session info

```{r}
sessionInfo()

```
